Structural analysis

### 2024-11-07 workflow/notebooks/draft.Rmd
implemented eggnog and oma benchmark

@@

dbs <- colnames(ortho_dbs)[2:dim(ortho_dbs)[2]]
# Loop through each column name and apply a function
jaccard_df <- map_dfr(dbs, ~ {
  col_name <- .x
select(df_filt, query, target, singleton) %>% 
    # filter(query=="AF-A0A024RBG1-F1") %>%
    left_join(select(ortho_dbs, target, col_name), by=c("query"="target")) %>% 
    left_join(select(ortho_dbs, target, col_name), by=c("target")) %>% 
    rename("query_set"=paste0(col_name, ".x"), "target_set"=paste0(col_name, ".y")) %>% 
    # Exclude queries without any domain
    filter(!is.na(query_set), !is.na(target_set)) %>% 
    group_by(query, singleton) %>%
    distinct(query, singleton, query_set, target_set) %>% 
    summarise(set_q = strsplit(paste(query_set, collapse = ""), ";"),
              set_t = strsplit(paste(target_set, collapse = ""), ";")) %>%
    rowwise() %>%
    mutate(jaccard = length(intersect(set_q, set_t))/length(union(set_q, set_t))) %>% 
    select(-set_q, -set_t) %>% 
    mutate(db=col_name)

@@

ggplot(jaccard_df, aes(jaccard, fill=singleton)) +
  geom_histogram(position = "dodge") +
  facet_wrap(~db, scales = "free_y") +
  scale_y_continuous(expand = expansion(0.01,0)) + 
  scale_fill_manual(values=palette_singleton_p) +
  theme_bw() +
  theme(legend.position = "top",
        # panel.background = element_rect(fill=NA, color="black"),
        # axis.line = element_blank()
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(), 
        strip.background = element_rect(fill=NA)
        # panel.grid.major.x = element_line()
        )

ggsave("test/bench.pdf", width = 8, height = 5, device = cairo_pdf)
# plot_dbs_list[[db]] <- p_db

>>>>

# ortho_dbs %>% 
#   left_join(taxmap, by=c("target"="label")) %>% 
#   group_by(mnemo) %>% 
#   summarise_all(~ sum(is.na(.))/nrow(ortho_dbs)) %>% 
#   filter(!is.na(mnemo)) %>% 
#   select(-target) %>% 
#   pivot_longer(!mnemo) %>% 
#   ggplot(aes(name, value)) + 
#   geom_boxplot() + 
#   labs(x="DB", y="% NA")

@@

dbs <- colnames(ortho_dbs)[2:dim(ortho_dbs)[2]]
dbs <- setdiff(dbs, c("InParanoid"))
dbs <- dbs[1:5]

binwidth <- 0.1

# Loop through each column name and apply a function
jaccard_df_mono <- map_dfr(dbs, ~ {
  col_name <- .x
  select(df_filt, query, target, singleton) %>% 
    left_join(select(ortho_dbs, target, col_name), by=c("query"="target")) %>% 
    left_join(select(ortho_dbs, target, col_name), by=c("target")) %>% 
    rename("query_set"=paste0(col_name, ".x"), "target_set"=paste0(col_name, ".y")) %>% 
    mutate(n_domain_q = str_count(pattern = "\\;", query_set),
           n_domain_t = str_count(pattern = "\\;", target_set)) %>% 
    filter(n_domain_q==1) %>% 
    # Exclude queries without any domain
    filter(!is.na(query_set), !is.na(target_set)) %>% 
    mutate(is_same = query_set==target_set) %>% 
    group_by(query, singleton) %>%
    summarise(jaccard=sum(is_same)/n()) %>% 
    # select(-set_q, -set_t) %>% 
    mutate(db=col_name)

@@

p_dbs_mono <- ggplot(jaccard_df_mono, aes(jaccard, fill=singleton)) +
  geom_histogram(aes(group=singleton, y=after_stat(binwidth*density)), 
                 position = "dodge", binwidth = binwidth) +
  facet_wrap(~db, ncol = 5) +
  scale_y_continuous(expand = expansion(0.01,0)) + 
  scale_fill_manual(values=palette_singleton_p) +
  theme_bw(base_family = "Aptos") +
  labs(title="Only monodomain", x="Jaccard Index", y="Scaled density", fill="Target sets") +
  theme(legend.position = "none",
        strip.background = element_rect(fill=NA))

# Loop through each column name and apply a function
jaccard_df_NA <- map_dfr(dbs, ~ {
  col_name <- .x
  select(df_filt, query, target, singleton) %>% 
    # filter(query=="AF-A0A024RBG1-F1") %>%
    left_join(select(ortho_dbs, target, col_name), by=c("query"="target")) %>% 
    left_join(select(ortho_dbs, target, col_name), by=c("target")) %>% 
    rename("query_set"=paste0(col_name, ".x"), "target_set"=paste0(col_name, ".y")) %>% 
    # Exclude queries without any domain
    filter(!is.na(query_set)) %>% 
    group_by(query, singleton) %>%
    distinct(query, singleton, query_set, target_set) %>% 
    summarise(set_q = strsplit(paste(query_set, collapse = ""), ";"),
              set_t = strsplit(paste(target_set, collapse = ""), ";")) %>%
    rowwise() %>%
    mutate(jaccard = length(intersect(set_q, set_t))/length(union(set_q, set_t))) %>% 
    select(-set_q, -set_t) %>% 
    mutate(db=col_name)

p_dbs_NA <- ggplot(jaccard_df_NA, aes(jaccard, fill=singleton)) +
  geom_histogram(aes(group=singleton, y=after_stat(binwidth*density)), 
                 position = "dodge", binwidth = binwidth) +
  facet_wrap(~db, ncol = 5) +
  scale_y_continuous(expand = expansion(0.01,0)) + 
  scale_fill_manual(values=palette_singleton_p) +
  theme_bw(base_family = "Aptos") +
  labs(title="NA=mismatch", x="Jaccard Index", y="Scaled density", fill="Target sets") +
  theme(legend.position = "none",
        strip.background = element_rect(fill=NA))

jaccard_df <- map_dfr(dbs, ~ {
  col_name <- .x
  select(df_filt, query, target, singleton) %>% 
    # filter(query=="AF-A0A024RBG1-F1") %>%
    left_join(select(ortho_dbs, target, col_name), by=c("query"="target")) %>% 
    left_join(select(ortho_dbs, target, col_name), by=c("target")) %>% 
    rename("query_set"=paste0(col_name, ".x"), "target_set"=paste0(col_name, ".y")) %>% 
    # Exclude queries without any domain
    filter(!is.na(query_set), !is.na(target_set)) %>% 
    group_by(query, singleton) %>%
    distinct(query, singleton, query_set, target_set) %>% 
    summarise(set_q = strsplit(paste(query_set, collapse = ""), ";"),
              set_t = strsplit(paste(target_set, collapse = ""), ";")) %>%
    rowwise() %>%
    mutate(jaccard = length(intersect(set_q, set_t))/length(union(set_q, set_t))) %>% 
    select(-set_q, -set_t) %>% 
    mutate(db=col_name)

p_dbs <- ggplot(jaccard_df, aes(jaccard, fill=singleton)) +
  geom_histogram(aes(group=singleton, y=after_stat(binwidth*density)), 
                 position = "dodge", binwidth = binwidth) +
  facet_wrap(~db, ncol = 5) +
  scale_y_continuous(expand = expansion(0.01,0)) + 
  scale_fill_manual(values=palette_singleton_p) +
  theme_bw(base_family = "Aptos") +
  labs(title="Default", x="Jaccard Index", y="Scaled density", fill="Target sets") +
  theme(legend.position = "top",
        strip.background = element_rect(fill=NA))

p_dbs / p_dbs_NA / p_dbs_mono

# ggsave("test/bench.pdf", p_dbs, width = 8, height = 5, device = cairo_pdf)
